---
title: "simpleRegress"
author: "April Parry"
date: "7/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

The Bears neck is 20 inches around which helps predicts about 57 in length. Looking at the accompanying plot, it shows that the residual standard error is 5.376, and the R-squared is .7524.


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 



```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```

The residual standard error is 4.939 and its lower than the previous model which is 5.376 and is better. In addition, the R-squared is 0.791, which is higher than the previous model, which is 0.7524 and also is better. So in ranking , chest is above Neck so we can conclude that chest is higher and better than the neck. 


```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```

As the Third model, the residual standard error is 7.32, which is making it worse. In addition, the the R-square is now 0.5409 which is also a decline from the previous model. Comparing all three models, This one in particular is at the bottom of the list. 



```{r}
MYWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)
```

This model is a comparison to the bears weight. The residual standard error is 5.432, and the adjusted R- square is 0.7472. In this case it is higher than the last model which is worse , and the adjusted R-square is higher than the previous model which is better. So whereas, the resiual standard error is worse, the adjusted R- sqaure is better. At this point the bear's chest has the highest adjusted R-square, which is the best. As far as the standard residual error

```{r}
MYAGE <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(MYAGE)
summary.lm(MYAGE)
```

The residual error is high with a 7.511 which is not a good result. While the R- squared is at 0.5166, Which is so far the best one out of all the plot charts we have doned. The smaller the R- squared, the better the results. 



```{r}
MYHEADLEN <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(MYHEADLEN)
summary.lm(MYHEADLEN)
```


The residual standard is at 4.235 on 52 degrees which is lower than the other chart above. While the R- squared for this data is 0.8463, which is pretty higher than the MYAGE plot and other plots we have doned. 
 Therefore, in comparison to the last model standard residual, this one is lower, therefore, it is better.

#CONCLUSION

In conclusion, when ranking against the bears length from his age, headlen, headwth, neck, chest and weight we have selected the order from best to worst due to their residual standard error: 

HEADLEN 4.235  
CHEST 4.939  
NECK 5.376  
WEIGHT 5.432  
HEADWTH 7.32  
AGE 7.511  


I learned how to plot data charts/graphs and determine each residual and their differences by spectating if the R-Squared is higher or lower than others. Also learned that the smaller the R-squared the better the results and the bigger the number the bad results we get. Furthermore I think the MYHEADLEN plot is quite larger than others due to its R-Squared which is making the results worse
